No meaningful effects of COVID-19 related social media use on well-being

### 2021-11-05 analyses.RmD
- Status of paper when submitted to ICA

```{r knitr-setup, include=F}
set.seed(170819)
knitr::opts_chunk$set(warning=F, echo=T, error=F, message=F, cache = T, 
                      tidy=T, dev = 'pdf', cache.extra = knitr::rand_seed)
options(digits = 3, width = 140, scipen=999)
```

@@

# define packages
packages <- c("brms", "corrplot", "devtools", "GGally", "ggplot2", 
              "gridExtra", "haven", "kableExtra", "knitr", 
              "lavaan", "lme4", "magrittr", "mice", "MVN", "PerFit", 
              "psych", "quanteda.textstats",
              "pwr", "semPlot", "semTools", "tidyverse")

@@

Find Variables in dataset.

@@

@@

@@

@@

>>>>

```{r knitr-setup, include=F}
set.seed(170819)
knitr::opts_chunk$set(warning=F, echo=T, error=F, message=F, cache = T, 
                      tidy=T, dev = 'pdf', cache.extra = knitr::rand_seed)
options(digits = 3, width = 200, scipen=999)
```

@@

# define packages
packages <- c("brms", "corrplot", "devtools", "GGally", "ggplot2", 
              "gridExtra", "haven", "kableExtra", "knitr", 
              "lavaan", "lme4", "magrittr", "mice", "MVN", "PerFit", 
              "psych", "quanteda.textstats",
              "pwr", "purrr", "semPlot", "semTools", "tidyverse")

@@

Find variables in dataset.

@@

Determine average reliability for measures across waves

```{r}
get_rel <- function(data, waves=24){

  # extract average reliability from lavaan fitted cfa with several groups
  reliability(data) %>% 
  unlist() %>% 
  matrix(5, waves) %>% 
  as.data.frame() %>% 
  set_rownames(c("alpha", "omega", "omega2", "omega3", "avevar")) %>% 
  summarise(omega = rowMeans(.["omega",])) %>% 
  return()
}
```

@@

Positive affect was measured as a scale. We hence first inspect factorial validty using CFA.

```{r}
model <- "
aff_pos =~ a1*aff_pos_1 + a2*aff_pos_2 + a3*aff_pos_3
"
cfa_aff_pos <- cfa(model, d_long, group = "wave")
summary(cfa_aff_pos, standardized = TRUE, fit = TRUE, estimates = FALSE)
```

The data fit the model very well, `td::fit_txt(cfa_aff_pos)`.
Let's next inspect reliability

```{r}
rel_aff_pos <- get_rel(cfa_aff_pos)
rel_aff_pos
```

The average reliability across all waves was omega = `r rel_aff_pos`, hence good.

Let's next expect the development across waves.

@@

Negative affect was measured as a scale. We hence first inspect factorial validty using CFA.

```{r}
model <- "
aff_neg =~ a1*aff_neg_1 + a2*aff_neg_2 + a3*aff_neg_3 + a4*aff_neg_4 + a5*aff_neg_5 + a6*aff_neg_6
"
cfa_aff_neg <- cfa(model, d_long, group = "wave")
summary(cfa_aff_neg, standardized = TRUE, fit = TRUE, estimates = FALSE)
```

The data fit the model very well, `td::fit_txt(cfa_aff_neg)`.

Let's next inspect reliability.

```{r}
rel_aff_neg <- get_rel(cfa_aff_neg)
rel_aff_neg
```

The average reliability across all waves was omega = `r rel_aff_neg`, hence good.

@@

## Locus of control

The only other variable that was measured as a scale was Locus of Control. Below I hence report the scale's factorial validity.

```{r}
model <- "
loc_cntrl_int =~ a1*loc_cntrl_int_1 + a2*loc_cntrl_int_2 + a3*loc_cntrl_int_3 + a4*loc_cntrl_int_4
# loc_cntrl_int_1 ~~ loc_cntrl_int_2
loc_cntrl_int_3 ~~ loc_cntrl_int_4
"
cfa_loc_cntrl_int <- cfa(model, filter(d_long, wave != 20), group = "wave")
summary(cfa_loc_cntrl_int, standardized = TRUE, fit = TRUE, estimates = FALSE)
```

The data fit the model very well, `td::fit_txt(cfa_loc_cntrl_int)`.

### 2021-11-05 index.Rmd
- Status of paper when submitted to ICA

---
title: "No effets of COVID-related social media use and well-being"
---

On this website, you can find the companion material for the paper _No effets of COVID-related social media use and well-being_.

>>>>

---
title: "No meaningful effets of COVID-related social media use and well-being"
---

On this website, you can find the companion material for the paper _No meaningful effets of COVID-related social media use and well-being_.

### 2021-11-05 manuscript.Rmd
- Status of paper when submitted to ICA

documentclass     : "apa6"
classoption       : "man"
# output            : papaja::apa6_pdf
# output            : papaja::apa6_word
header-includes:
  - \setlength{\parskip}{0em}
  - \raggedbottom
  - \note{\clearpage}
---

@@

Several people reported that they were glued to their screens and could not pursue other relevant activities such as working, taking a break, or even child-care [@kleinDarklySoothingCompulsion2021].
As doomscrolling increased it became doubtful whether such a surge in social media use could still be considered useful and adaptive, or whether it created an additional and new psychological danger for users' mental health [@sandstromDoomscrollingCOVIDNews2021].
These concerns seem justified: A study with 6,233 people from Germany found that "[f]requency, duration and diversity of media exposure were positively associated with more symptoms of depression and unspecific and COVID-19 specific anxiety" [@bendauAssociationsCOVID19Related2021], which implies negative consequences for well-being.

@@

In this study, life satisfaction was measured on an 11-point scale, which is more than the seven degrees people can reliably differentiate. 
We hence need to transform a 1-point change on a 7-point scale to its equivalent on an 11-point scale, which is 11 / 7 = `r 11/7`.
Hence, we now expect our effect to be at least as large as _b_ = .25 * `r 11/7` = `r 11/7*.25`.
For affect, which was measured on a 5-point scale, our SESOI would therefore be _b_ = 1/4 * 5/7 = `r 5 / 7 * .25`.
Because we are agnostic as to whether the effects are positively or negatively nontrivial, the null region will include negative and positive effects (in Bayesian terms, this is called the region of practical equivalence [ROPE]).
In order not to exaggerate precision, and to be a bit less conservative, these numbers will be reduced to nearby thresholds. 
Note that other researchers also decreased or recommended decreasing thresholds for effect sizes when analyzing within-person or cumulative effects (@beyensSocialMediaUse2021; @funderEvaluatingEffectSize2019).
Together, this leads to an indifference region ranging from _b_ = -.30 to _b_ = .30 for life satisfaction, and _b_ = -.15 to _b_ = .15 for positive and negative affect.

@@

To capture positive affect, respondents were asked how often in the last week they felt (a) calm and relaxed, (b) happy, and (c) full of energy.
The response options were 1 (_never_), 2 (_on some days_), 3 (_several times per week_), 4 (_almost every day_), and 5 (_daily_).

For negative affect, respondents were asked how often in the last week they felt (a) lonely, (b) aggravated, (c) so depressed, that nothing could lift you up, (d) very nervous, (e) anxious, and (h) glum and sad.
The response options were 1 (_never_), 2 (_on some days_), 3 (_several times per week_), 4 (_almost every day_), and 5 (_daily_).

>>>>

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
# output            : papaja::apa6_word
header-includes:
  - \setlength{\parskip}{0em}
  - \raggedbottom
  - \note{\clearpage}
---

@@

Several people reported that they were glued to their screens and could not pursue other relevant activities such as working, taking a break, or even child-care [@kleinDarklySoothingCompulsion2021].
As doomscrolling increased it became doubtful whether such a surge in social media use could still be considered useful and adaptive, or whether it created an additional hazard for users' mental health [@sandstromDoomscrollingCOVIDNews2021].
These concerns seem justified: A study with 6,233 people from Germany found that "[f]requency, duration and diversity of media exposure were positively associated with more symptoms of depression and unspecific and COVID-19 specific anxiety" [@bendauAssociationsCOVID19Related2021], which implies negative consequences for well-being.

@@

In this study, life satisfaction was measured on an 11-point scale, which is more than the seven degrees people can reliably differentiate. 
We hence need to transform a 1-point change on a 7-point scale to its equivalent on an 11-point scale, which is 11 / 7 = `r 11/7`.
Hence, we now expect our effect to be at least as large as _b_ = .25 * `r 11/7` = `r 11/7*.25`.
For affect, which was measured on a 5-point scale, our SESOI would therefore be _b_ = 1/4 * 5/7 = `r 5 / 7 * .25`.
Because we are agnostic as to whether the effects are positively or negatively nontrivial, the null region will include negative and positive effects (in Bayesian terms, this is called the region of practical equivalence [ROPE]).
In order not to exaggerate precision, and to be a bit less conservative, these numbers will be reduced to nearby thresholds. 
Note that other researchers also decreased or recommended decreasing thresholds for effect sizes when analyzing within-person or cumulative effects [@beyensSocialMediaUse2021; @funderEvaluatingEffectSize2019].
Together, this leads to an indifference region ranging from _b_ = -.30 to _b_ = .30 for life satisfaction, and _b_ = -.15 to _b_ = .15 for positive and negative affect.

@@

To capture positive affect, respondents were asked how often in the last week they felt (a) calm and relaxed, (b) happy, and (c) full of energy.
The response options were 1 (_never_), 2 (_on some days_), 3 (_several times per week_), 4 (_almost every day_), and 5 (_daily_).
The scale showed good factorial fit, `r fit_txt(cfa_aff_pos)`.
Reliability was also very high, omega = `r rel_aff_pos %>% my_round("std")`.

For negative affect, respondents were asked how often in the last week they felt (a) lonely, (b) aggravated, (c) so depressed, that nothing could lift you up, (d) very nervous, (e) anxious, and (h) glum and sad.
The response options were 1 (_never_), 2 (_on some days_), 3 (_several times per week_), 4 (_almost every day_), and 5 (_daily_).
The scale showed good factorial fit, `r fit_txt(cfa_aff_neg)`.
Reliability was also very high, omega = `r rel_aff_neg %>% my_round("std")`.

### 2021-11-05 manuscript.tex
- Status of paper when submitted to ICA

Several people reported that they were glued to their screens and could not pursue other relevant activities such as working, taking a break, or even child-care (Klein, 2021).
As doomscrolling increased it became doubtful whether such a surge in social media use could still be considered useful and adaptive, or whether it created an additional and new psychological danger for users' mental health (Sandstrom, Buchanan, Aknin, \& Lotun, 2021).
These concerns seem justified: A study with 6,233 people from Germany found that ``{[}f{]}requency, duration and diversity of media exposure were positively associated with more symptoms of depression and unspecific and COVID-19 specific anxiety'' (Bendau et al., 2021), which implies negative consequences for well-being.

@@

In this study, life satisfaction was measured on an 11-point scale, which is more than the seven degrees people can reliably differentiate.
We hence need to transform a 1-point change on a 7-point scale to its equivalent on an 11-point scale, which is 11 / 7 = 1.57.
Hence, we now expect our effect to be at least as large as \emph{b} = .25 * 1.57 = 0.39.
For affect, which was measured on a 5-point scale, our SESOI would therefore be \emph{b} = 1/4 * 5/7 = 0.18.
Because we are agnostic as to whether the effects are positively or negatively nontrivial, the null region will include negative and positive effects (in Bayesian terms, this is called the region of practical equivalence {[}ROPE{]}).
In order not to exaggerate precision, and to be a bit less conservative, these numbers will be reduced to nearby thresholds.
Note that other researchers also decreased or recommended decreasing thresholds for effect sizes when analyzing within-person or cumulative effects (Beyens, Pouwels, Driel, Keijsers, and Valkenburg (2021); Funder and Ozer (2019)).
Together, this leads to an indifference region ranging from \emph{b} = -.30 to \emph{b} = .30 for life satisfaction, and \emph{b} = -.15 to \emph{b} = .15 for positive and negative affect.

@@

To capture positive affect, respondents were asked how often in the last week they felt (a) calm and relaxed, (b) happy, and (c) full of energy.
The response options were 1 (\emph{never}), 2 (\emph{on some days}), 3 (\emph{several times per week}), 4 (\emph{almost every day}), and 5 (\emph{daily}).

For negative affect, respondents were asked how often in the last week they felt (a) lonely, (b) aggravated, (c) so depressed, that nothing could lift you up, (d) very nervous, (e) anxious, and (h) glum and sad.
The response options were 1 (\emph{never}), 2 (\emph{on some days}), 3 (\emph{several times per week}), 4 (\emph{almost every day}), and 5 (\emph{daily}).

>>>>

Several people reported that they were glued to their screens and could not pursue other relevant activities such as working, taking a break, or even child-care (Klein, 2021).
As doomscrolling increased it became doubtful whether such a surge in social media use could still be considered useful and adaptive, or whether it created an additional hazard for users' mental health (Sandstrom, Buchanan, Aknin, \& Lotun, 2021).
These concerns seem justified: A study with 6,233 people from Germany found that ``{[}f{]}requency, duration and diversity of media exposure were positively associated with more symptoms of depression and unspecific and COVID-19 specific anxiety'' (Bendau et al., 2021), which implies negative consequences for well-being.

@@

In this study, life satisfaction was measured on an 11-point scale, which is more than the seven degrees people can reliably differentiate.
We hence need to transform a 1-point change on a 7-point scale to its equivalent on an 11-point scale, which is 11 / 7 = 1.57.
Hence, we now expect our effect to be at least as large as \emph{b} = .25 * 1.57 = 0.39.
For affect, which was measured on a 5-point scale, our SESOI would therefore be \emph{b} = 1/4 * 5/7 = 0.18.
Because we are agnostic as to whether the effects are positively or negatively nontrivial, the null region will include negative and positive effects (in Bayesian terms, this is called the region of practical equivalence {[}ROPE{]}).
In order not to exaggerate precision, and to be a bit less conservative, these numbers will be reduced to nearby thresholds.
Note that other researchers also decreased or recommended decreasing thresholds for effect sizes when analyzing within-person or cumulative effects (Beyens, Pouwels, Driel, Keijsers, \& Valkenburg, 2021; Funder \& Ozer, 2019).
Together, this leads to an indifference region ranging from \emph{b} = -.30 to \emph{b} = .30 for life satisfaction, and \emph{b} = -.15 to \emph{b} = .15 for positive and negative affect.

@@

To capture positive affect, respondents were asked how often in the last week they felt (a) calm and relaxed, (b) happy, and (c) full of energy.
The response options were 1 (\emph{never}), 2 (\emph{on some days}), 3 (\emph{several times per week}), 4 (\emph{almost every day}), and 5 (\emph{daily}).
The scale showed good factorial fit, \(\chi^2\)(46) = 65.30, \textit{p} = .032, cfi = 1.00, rmsea = .02, 90\% CI {[}.01, .03{]}, srmr = .01.
Reliability was also very high, omega = .85.

For negative affect, respondents were asked how often in the last week they felt (a) lonely, (b) aggravated, (c) so depressed, that nothing could lift you up, (d) very nervous, (e) anxious, and (h) glum and sad.
The response options were 1 (\emph{never}), 2 (\emph{on some days}), 3 (\emph{several times per week}), 4 (\emph{almost every day}), and 5 (\emph{daily}).
The scale showed good factorial fit, \(\chi^2\)(331) = 3138.37, \textit{p} \textless{} .001, cfi = .97, rmsea = .08, 90\% CI {[}.07, .08{]}, srmr = .03.
Reliability was also very high, omega = .89.
